---
title: "Empirical Project 1"
author: "Ben Noon"
format: html
editor: visual
---

## Load Library and Data

```{r}
library(pacman)
p_load(tidyverse, stargazer, data.table, tidyfast, dtplyr, haven)

ohp = read_dta("ohp.dta")
ohp_dt <- as.data.table(ohp)
```

# Questions

### Q1

Explain the difference between the variables treatment and ohp_all_ever_survey. Explain\
why treatment is the treatment variable (Di), rather than ohp_all_ever_survey.

Treatment is when someone either won the OHP lottery or did not win the lottery (1, 0).

ohp_all_ever_survey is if the person was every enrolled into medicaid.

Treatment is the variable we want because we are looking to see if increased access to medicaid improves health outcomes. ohp_all_ever_survey is independent of the research question and gives us a look at the compliance rate.

### Q2

Provide evidence that the OHP lottery really did randomly assign individuals to treatment\
and control groups. Similar to Table 1 in Taubman et al. (2014), please create a nicely\
formatted table that reports means of 4 to 6 relevant characteristics for individuals in the\
control group

```{r}

edu <- lm(data = ohp_dt, edu_inp ~ treatment)
age <- lm(data = ohp_dt, age_inp ~ treatment)
med_ever <- lm(data = ohp_dt, ohp_all_ever_survey ~ treatment)
black <- lm(data = ohp_dt, race_black_inp~treatment)
white <- lm(data = ohp_dt, race_white_inp~treatment)
other <- lm(data = ohp_dt, race_nwother_inp~treatment)
dep <- lm(data = ohp_dt, dep_dx_pre_lottery ~ treatment)
dia <- lm(data = ohp_dt, dia_dx_pre_lottery ~ treatment)

stargazer(edu, age, med_ever, black, white, other, dep, dia, type="html")
```
